World Heritage Sites on Wikipedia: Inscription Controversies in a Context of Constrained Agency for Cultural Heritage Activism

### 2020-05-26 analysis/paper/999-code-for-the-Stirling-ppt.Rmd
start to write results

# load the random pages
ten_k_random_wp_pages <- readRDS(here::here("data", "random_page_data_tbl.rds"))

# load the random pages revision histories
ten_k_random_wp_pages_lst <- readRDS(here::here("data", "random_page_data_lst.rds"))

# all WH sites with WP pages (n = 584)
page_data_for_all_pages <-
  readRDS(here::here("data",
                     "page_data_for_all_pages.rds"))

# all WH sites, with and without WP pages (n = 846)
wh_wiki_table <-
  read_csv(here::here("data",
                      "wh_wiki_table.csv"))

# the talk pages for all WH sites, with WP pages (n = 846)
wh_wiki_talk_pages <-  readRDS(here::here("data",
                     "wh_wiki_table_talk_page_content.rds"))

# all WH sites according to UNESCO
# from https://whc.unesco.org/en/syndication
wh_unesco <- readxl::read_excel(here::here("data",
                                           "whc-sites-2018.xls")) %>%
  filter(category %in% c("Cultural"))

>>>>

# load the random pages
ten_k_random_wp_pages <- readRDS(here::here("analysis/data/raw_data/random_page_data_tbl.rds"))

# load the random pages revision histories
ten_k_random_wp_pages_lst <- readRDS(here::here("analysis/data/raw_data/random_page_data_lst.rds"))

# all WH sites with WP pages (n = 584)
page_data_for_all_pages <-
  readRDS(here::here("analysis/data/raw_data/page_data_for_all_pages.rds"))

# all WH sites, with and without WP pages (n = 846)
wh_wiki_table <-
  readxl::read_excel(here::here("analysis/data/raw_data/whc-sites-2019.xls"))

# the talk pages for all WH sites, with WP pages (n = 846)
wh_wiki_talk_pages <-  readRDS(here::here("analysis/data/raw_data/wh_wiki_table_talk_page_content.rds"))

# all WH sites according to UNESCO
# from https://whc.unesco.org/en/syndication
wh_unesco <- wh_wiki_table %>%
  filter(category %in% c("Cultural"))

### 2020-05-26 analysis/paper/paper.Rmd
start to write results

Several CS-WHL sites are notable for the conflicts and tensions that have surrounded their inscription. The 1992 inscription of Angkor (an ancient city and empire in Cambodia, prominent during the 9th to the 15th centuries AD) was supported by exiled supporters of the genocidal Khmer Rouge regime, hoping to strengthen territorial claims [@locard2015myth]. They appropriated Western discourse on national cultural heritage to argue for the safeguarding of Angkor as part of their quest for national independence and international recognition. Early in the Khmer Rouge regime Angkor was declared a symbol of enslavement by a primitive culture, but when the Khmer Rouge adopted a new rhetoric of a supposedly civilizing mission, they presented it as the site one of the great world civilizations [@falser2015representing]. The 2003 inscription of Mapungubwe (the site of the first indigenous kingdom in Southern Africa, 900-1,300 AD) was preceded by a recommendation from ICOMOS (International Council on Monuments and Sites, a professional association that is a key advisory body to the Wold Heritage Committee) not to inscribe because of the farming and mining activity in highly sensitive areas near the site, and the unclear ownership of the mining rights at the time [@meskell2011paris]. Despite this negative recommendation, geopolitical machinations within the Committee, especially by the Indian and Russian delegates, lead to Mapungubwe being inscribed on the list, although without the typical prerequisites of an integrated management plan or complete buffer zone [@meskell2012rush]. These examples of Angkor and Mapungubwe demonstrate the attention that the WHL inscription process can generate due to political activitsm, conflicts and intrigue.

Physical conflict at or near CS-WHL are also major events that galvanise public interest in these locations. World Heritage sites in Palestine, Mali, Syria, Congo and Cambodia have recently been sites of violence, in many cases the violence is specifically linked to their potential WHL nomination, listing or management. In 1998 anti-government and mostly Hindu Tamil groups bombed the holy Buddhist site of the Temple of the Tooth at the WHL site of Kandy (the last capital of the ancient kings of Sri Lanka), killing 17 people and substantially damaging the temple [@coningham1999paradise]. In Mali during 2012, fighting between government and rebel groups lead to the damage and destruction of tombs at the CS-WHL sites of Gao and Timbuktu [@brioschi2017discourses]. The World Heritage Committee found itself powerless to intervene because of political gridlock [@meskell2015gridlock], and these Mali sites are currently among the 53 sites on the List of World Heritage in Danger (https://whc.unesco.org/en/danger/). In 2015 ISIS militants destroyed the Temple of Bel in Palmyra, Syria (a CS-WML site of monumental ruins, once great city at the crossroads between east and west in the ancient world) [@gornik2015can]

in Kandy (Sri Lanka) ,
Palmyra (Syria), etc.
• Preah Vihear (2008) on a disputed
part of Thai-Cambodia border,
WikiLeaks documents the
hyperconnectivity of international
actors

@@

We survey the basic characteristics of content (article length, number of wikilinks out to other pages, number of citations to non-Wikipedia items), consumption (page view counts, wikilinks in from other Wikipedia pages), and production (edit counts, edit densities, edit sizes, number of unique editors per article, talk page length, talk page topics) to answer the question: how does engagement with the past compare on Wikipedia compare to engagement with other topics? We use a random sample of 10,000 Wikipedia articles as a random sample to compare with articles about cultural sites on the UNESCO World Heritage List. We investigate spatial patterns of behaviour on these Wikipedia articles to answer the question:  are the editors of CS-WHL located near the sites they edit, indicating local community interest in the online representation of their heritage, or in other countries, impling a digital colonalism of world heritage information on Wikipedia where one community's heritage is interpreted and communicated by another? Finally we investigate temporal patterns of activity on CS-WHL Wikipedia pages to answer the question of how correlated editing activity is with events outside of Wikipedia relating to the CS-WHL sites, such as their inscription on the WHL.

In this paper we contribute to digital heritage studies into a new online domain

@@

@@

<!-- The following line inserts a page break when the output is MS Word. For page breaks in PDF, use \newpage on its own line.  -->
##### pagebreak

@@

Kristiansen, K. 2014. “Towards a New Paradigm? the Third Science Revolution and Its Possible Consequences in Archaeology.” Current Swedish Archaeology 22: 11–34.

##### pagebreak

>>>>

Several CS-WHL sites are notable for the conflicts and tensions that have surrounded their inscription. The 1992 inscription of Angkor (an ancient city and empire in Cambodia, prominent during the 9th to the 15th centuries AD) was supported by exiled supporters of the genocidal Khmer Rouge regime, hoping to strengthen territorial claims [@locard2015myth]. They appropriated Western discourse on national cultural heritage to argue for the safeguarding of Angkor as part of their quest for national independence and international recognition. Early in the Khmer Rouge regime Angkor was declared a symbol of enslavement by a primitive culture, but when the Khmer Rouge adopted a new rhetoric of a supposedly civilizing mission, they presented it as the site one of the great world civilizations [@falser2015representing]. The 2003 inscription of Mapungubwe (the site of the first indigenous kingdom in Southern Africa, 900-1,300 AD) was preceded by a recommendation from ICOMOS (International Council on Monuments and Sites, a professional association that is a key advisory body to the World Heritage Committee) not to inscribe because of the farming and mining activity in highly sensitive areas near the site, and the unclear ownership of the mining rights at the time [@meskell2011paris]. Despite this negative recommendation, geopolitical machinations within the Committee, especially by the Indian and Russian delegates, led to Mapungubwe being inscribed on the list, although without the typical prerequisites of an integrated management plan or complete buffer zone [@meskell2012rush]. These examples of Angkor and Mapungubwe demonstrate the attention that the WHL inscription process can generate due to political activitsm, conflicts and intrigue.

Physical conflict at or near CS-WHL are also major events that galvanise public interest in these locations. World Heritage sites in Palestine, Mali, Syria, Congo and Cambodia have recently been sites of violence, in many cases the violence is specifically linked to their potential WHL nomination, listing or management. In 1998 anti-government and mostly Hindu Tamil groups bombed the holy Buddhist site of the Temple of the Tooth at the WHL site of Kandy (the last capital of the ancient kings of Sri Lanka), killing 17 people and substantially damaging the temple [@coningham1999paradise]. In Mali during 2012, fighting between government and rebel groups lead to the damage and destruction of tombs at the CS-WHL sites of Gao and Timbuktu [@brioschi2017discourses]. The World Heritage Committee found itself powerless to intervene because of political gridlock [@meskell2015gridlock], and these Mali sites are currently among the 53 sites on the List of World Heritage in Danger (https://whc.unesco.org/en/danger/). In 2015, ISIS militants destroyed the Temple of Bel in Palmyra, Syria (a CS-WML site of monumental ruins, once great city at the crossroads between east and west in the ancient world) [@gornik2015can]. Preah Vihear, inscribed in 2008, is a CS-WHL located on a long-disputed section of the Thai-Cambodia border that has been a site of both violent military clashes and international political intrigue. Although both Thailand and Cambodia supported the nomination of the site to the WHL, the Thai government objected to maps in the nomination package that showed Cambodia as the owner of disputed land next to the temple, leading to protests and military clashes [@sothirak2013cambodia]. @meskell2020world's analysis of US diplomatic cables released by WikiLeaks further reveals that settlement of disputes over Preah Vihear were intricately tied to broader issues of foreign policy and US and Chinese investment, especially access to natural gas reserves in the Gulf of Thailand. For example, the cables reveal negotiations for Cambodia to retain their temple, leading to the enhancement of Thailand's underwater assets, and the United States negotiating for extended contracts with both countries.

@@

Our brief review of controversial cultural sites on the WHL shows the intensity and diversity of conflicts and tensions that surround these sites. Many CS-WHL are symbols of national, cultural, political, and religious identity, and the extent of political involvement in negotiations of WHL inscriptions indicates they are of great public interest among local and diasporic communities. Our goal in this study is answer the question of how this interest is expressed within the socio-technical constraints of Wikipedia. How are conflicts over CS-WHL enacted in the canalized interactions of the world's largest encyclopedia? We attempt to answer this question by studying all the CS-WHL that have Wikipedia articles. We surveyed the basic characteristics of content (article length, number of wikilinks out to other pages, number of citations to non-Wikipedia items), consumption (page view counts, wikilinks in from other Wikipedia pages), and production (edit counts, edit densities, edit sizes, number of unique editors per article, talk page length, talk page topics). By comparing these basic characteristics of Wikipedia articles about CS-WHL to 10,000 random Wikipedia articles we will answer the question: how does engagement with the past compare via CS-WHL on Wikipedia compare to engagement with other topics? Do Wikipedia articles about controversial CS-WHL sites, such as those discussed above, get more attention than random Wikipedia articles? Can we detect conflict in the edit histories, bot activity, and talk pages for Wikipedia articles about CS-WHL sites, and how does this conflict relate to the types of controversies noted above?

The highly detailed edit histories that Wikipedia keeps for every article allow us to further investigate spatial and temporal questions relating to engagement with the past and conflicts surrounding CS-WHL sites. For example, when an article is edited by a user who does not have a Wikipedia user account (or is not logged into their account), their edit is identified by that person's IP address. This IP address can be used to geo-locate the user to the city they were in when the made the edit. We geo-located all edits with IP addresses for all Wikipedia articles CS-WHL sites to determine the country of origin of those edits. This helps us to answer the question: are the editors of articles about CS-WHL located near the sites they edit, indicating local community interest in the online representation of their heritage? Or are editors located in other countries, implying a digital colonialism of world heritage information on Wikipedia where one community's heritage is interpreted and communicated by another? The time and date stamps attached to every edit on every article allows us to investigate temporal patterns of activity on CS-WHL Wikipedia articles. Analyses of these data help us to answer the question: is Wikipedia editing activity correlated with events outside of Wikipedia relating to the CS-WHL sites, such as conflict events, or their inscription on the WHL?

We obtained data about Wikipedia articles by scraping the HTML pages with the rvest package for R [@rvest]. We used the SelectorGadget [@cantino2017selectorgadget] extension for the Chrome web browser to identify specific elements of interest, or nodes, on the HTML pages, and wrote custom R functions to extract data from these nodes. The highly structured nature of most Wikipedia pages afforded us much efficiency during data collection, however, a not insignificant number of pages had minor irregularities, for example, exotic table formatting or minor deviations from widely used templates, that necessitated more complex programming. An especially challenging issue was minor irregularities in spelling place names, for example, "Saint Kitts and Nevis" and "St. Kitts and Nevis", which can confound programmatic joining and summarizing of data by failing to recognize as the same place and leading to under-counting. We attempted to correct these manually in our code. We collected data during May 2019, and due to the dynamic nature of Wikipedia, it is highly likely that articles in our study have subtly changed since our data collection as users and bots continue their activity. It is unlikely that articles in our sample have been deleted from Wikipedia since our data collection, because of the high notability of CS-WHL, but it is possible that new articles have since appeared about sites that were not represented on Wikipedia since we collected the data.

# Reproducibility and Open Source Materials for This Study

To enable reuse of our materials and improve reproducibility and transparency [@marwick2017computational], we include the entire R code [@rprogram] used for all the analysis and visualizations contained in this article in our compendium at http://doi.org/10.17605/OSF.IO/AY27G. Also in this version-controlled compendium are the raw data for all the results reported here. All of the figures and quantitative results presented here can be independently reproduced with the code and data in this repository. In our compendium our code is released under the MIT license, our data as CC-0, and our figures as CC-BY, to enable maximum reuse (for more details, see @marwick2018packaging).

@@

```{r wh-wiki-map, fig.cap="Wikipedia articles for Cultural Sites on the UNESCO World Heritage List. Countries coloured black currently have no Wikipedia articles for CS-WHL sites. Inset shows the distribution of articles per country. Map data from naturalearthdata.com", fig.width=15}
source(here::here("analysis/paper/002-Wikipedia-CS-WHL-world-map-counts.R"))
knitr::include_graphics(here::here("analysis/figures/wh_country_wikipedia_count_map.png"))
```

Of the `r nrow(wh_unesco)` cultural sites on the World Heritage List at the time of writing, we found Wikipedia articles for `r sum(wh_wk_country_counts$n)`. To characterize the overall variability of article according to the basic characteristics described above, we applied a Uniform Manifold Approximation and Projection (UMAP) to reduce the dimensionality of our data. After a two-dimensional projection was obtained, we used Hierarchical Density-Based Spatial Clustering of Applications with Noise (HDBSCAN) to identify clusters, and Random Forests to identify variable loadings so we could determine which of the basic characteristics were most important in driving variability between the articles.

Figure \@ref(fig:wh-wiki-map) shows that most countries have between 2-3 CS-WHL on Wikipedia, with Germany, India, China, France and Spain being the countries with the most articles about their CS-WHL.

@@

Thanks to Chiara Bonacchi, Rodney Harrison, and Daniel Pett for inviting BM to present this work at the "Digital Heritage in a World of Big Data" conference at the University of Stirling, Scotland, May 2019. This conference was part of the AHRC-funded project 'Ancient Identities in Modern Britain' and the AHRC Heritage Priority Area Leadership Fellowship.

@@
